---
title: "Assignment"
author: "Signe G"
date: "2024-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(epiwraps)
  library(ggplot2)
  library(rGREAT) 
})
```


```{r, eval=FALSE}
options(timeout = 6000)
unzip("w10.assignment.zip")
list.files()
```

## Get the data

```{r}
tracks <- list.files(pattern="bw$")
peaks <- list.files(pattern="bed$")
# we first import the peaks
peaks <- lapply(peaks, rtracklayer::import.bed)
# we'll focus on the high-quality peaks
peaks <- lapply(peaks, FUN=function(x) x[x$score>800])
# we get the union of non-redundant regions
regions <- reduce(unlist(GRangesList(peaks)))
```

## Clustering

```{r}
ese <- signal2Matrix(tracks, regions, extend=2000)
set.seed(123)  # to ensure that it gives the same results everytime
cl <- clusterSignalMatrices(ese, k=4)
rowData(ese)$cluster <- cl
```

## Plot the clustering

```{r}
plotEnrichedHeatmaps(ese, row_split=cl,
                     colors=c("white","darkred"))
```

## Trying different numbers of clusters

```{r}
cl2 <- clusterSignalMatrices(ese, k=2:10)
ggplot(cl2$varExplained, aes(k, varExplained)) + geom_line()
```

## Plot using relative signal using 7 clusters

```{r}
cl <- clusterSignalMatrices(ese, k=7, scaleRows = TRUE)
rowData(ese)$cluster <- cl
d <- meltSignals(ese, splitBy=cl)
ggplot(d, aes(position, mean, colour=sample)) + geom_line() + facet_wrap(~split)
plotEnrichedHeatmaps(ese, row_split = cl, scale_rows = "global")
```


## Enrichment analysis

```{r}
split_regions <- split(rowRanges(ese), rowData(ese)$cluster)
res <- great(split_regions[["7"]], gene_sets="GO:BP", tss_source="hg38", 
             background=regions, cores=2)
bp <- getEnrichmentTables(res)
```

## We plot the top Biological Processes:

```{r, fig.width=9, fig.height=6}
ggplot(head(bp,15), aes(fold_enrichment, reorder(description, p_adjust), 
                        size=observed_region_hits, color=-log10(p_adjust))) + 
  geom_point() + scale_color_viridis_c()
```

## Results

We are observing different binding patterns for different regions. I have found 7 different binding patterns/clusters for the 3 TFs. The enrichment analysis shows that different biological processes are enriched, like cell differentiation, cellular developmental process, and cell projection assembly. 


