---
title: "Assignment"
author: "Signe G"
date: "2024-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages({
  library(epiwraps)
  library(motifmatchr)
  library(MotifDb)
  library(universalmotif)
  library(ggplot2)
  library(SummarizedExperiment)
  library(sechm)
  library(BiocParallel)
  library(chromVAR)
  library(limma)
})

register(SnowParam(2))

```

```{r, eval=FALSE}
options(timeout=6000)
download.file("https://ethz-ins.org/content/mouse_mm38_hippocampus.peakCounts.SE.rds", "mouseHC_peakCounts.rds", mode="wb")

```
## Get the database of motifs

```{r}
motifs <- query(MotifDb, c("HOCOMOCOv11-core", "Hsapiens"))
motifs <- do.call(TFBSTools::PWMatrixList, setNames(
           universalmotif::convert_motifs(motifs, class="TFBSTools-PWMatrix"),
           mcols(motifs)$geneSymbol))
```


## Get the genome
```{r}
library(BSgenome.Mmusculus.UCSC.mm10)
genome <- BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
seqlevelsStyle(genome) <- "Ensembl"
```


```{r}
se <- readRDS("mouseHC_peakCounts.rds")
colData(se)
se$condition <- c("Control_f", "Control_f", "Control_f", "Control_m", "Control_m", "Control_m", "FSS_f", "FSS_f", "FSS_f", "FSS_m", "FSS_m", "FSS_m")
row.names(se) <-  as.character(granges(se))

```

*GC concentration*:
```{r}
suppressWarnings(se <- chromVAR::addGCBias(se, genome=genome))
rowData(se)
```

```{r}
suppressWarnings(moi <- motifmatchr::matchMotifs(motifs, subject=se, genome=genome))
bg <- chromVAR::getBackgroundPeaks(se, niterations=1000)
dev <- chromVAR::computeDeviations(object = se, annotations=moi,
                                   background_peaks=bg)
dev$condition <- factor(dev$condition)
```


## Differential analysis for stressed vs control

```{r}
mm_stress_vs_control <- model.matrix(~ dev$condition)
fit_stress_vs_control <- eBayes(lmFit(assays(dev)$z, mm_stress_vs_control))
res_stress_vs_control <- as.data.frame(topTable(fit_stress_vs_control, coef = "dev$conditionFSS_f", number = Inf))

```

```{r}
res_stress_vs_control$TF <- row.names(res_stress_vs_control)
sechm(dev, features = head(row.names(res_stress_vs_control)), assayName="z", top_annotation = c("condition", "depth"))
```

*Interpretation*
We see that GCR and PRGR is upregulated in stressed mice compared to control mice. The rest of the motifs do not tell a clear story.

## Differential analysis for female vs male
```{r}
mm_male_vs_female <- model.matrix(~ dev$condition)
fit_male_vs_female <- eBayes(lmFit(assays(dev)$z, mm_male_vs_female))
res_male_vs_female <- as.data.frame(topTable(fit_male_vs_female, coef = "dev$conditionControl_m", number = Inf))
```

```{r}
res_male_vs_female$TF <- row.names(res_male_vs_female)
sechm(dev, features = head(row.names(res_male_vs_female)), assayName="z", top_annotation = c("condition", "depth"))
```

*Interpretation*
We see that the motifs ZFP42, TYY1, ZN384 especially, are upregulated in females and downregulated in males. 
